---
output:
  pdf_document: default
  html_document: default
---
```{r}
library(tidyverse)
library(brms)
library(bayesplot)
library(loo)

options(mc.cores = parallel::detectCores())
rstan::rstan_options(auto_write = TRUE)
```

```{r load_data}
df_p1 <- read.csv("/Users/debarpita/Desktop/arjun/trial_wise_dataset_post.csv")
df_p2 <- read.csv("/Users/debarpita/Desktop/arjun/trial_wise_dataset_pre.csv")

df <- bind_rows(df_p1, df_p2)
```


```{r}
df <- df %>%
  filter(!is.na(Reward), Reward != 0)
df <- df %>%
  group_by(Participant_ID, stim_cat, env) %>%
  arrange(patch_id, .by_group = TRUE) %>%
  mutate(patch_number = patch_id) %>%
  ungroup()

# Remove patches with zero cumulative reward
df <- df %>%
  group_by(Participant_ID, stim_cat, env, patch_number) %>%
  filter(min(TotalCumulativeReward, na.rm = TRUE) > 0) %>%
  ungroup()

# Z-score predictors
df <- df %>%
  mutate(
    patch_number_z = scale(patch_number)[,1],
    trait_anxiety_score_z = scale(trait_anxiety_score)[,1]
  )

summary(df$TotalCumulativeReward)
```

```{r}
df %>%
  group_by(stim_cat, env) %>%
  summarise(
    n = n(),
    min_cum = min(TotalCumulativeReward, na.rm = TRUE),
    max_cum = max(TotalCumulativeReward, na.rm = TRUE),
    mean_cum = mean(TotalCumulativeReward, na.rm = TRUE),
    .groups = "drop"
  )

```




Model 1: Learning Across Patches
Do participants increase cumulative reward across patches within each environment, and does stimulation alter the rate of reward growth?
Because cumulative reward is strictly positive and right-skewed, we use a Gamma regression with log link.

```{r}
model1 <- brm(
  TotalCumulativeReward ~ patch_number_z * stim_cat + env +
    (1 | Participant_ID),
  data = df,
  family = Gamma(link = "log"),
  chains = 4,
  iter = 4000,
  warmup = 1500,
  cores = 4,
  seed = 123
)
```

```{r}
summary(model1)
```

beta_patch_number_z > 0:  Cumulative reward increases across patches  4% increase

```{r}
exp(fixef(model1))
```

```{r}
pp_check(model1, type = "dens_overlay", ndraws = 100)
```

```{r}
pp_check(model1, type = "stat_grouped", group = "stim_cat")
```

```{r}
model2 <- brm(
  TotalCumulativeReward ~ patch_number_z * stim_cat *
    trait_anxiety_score_z + env +
    (1  | Participant_ID),
  data = df,
  family = Gamma(link = "log"),
  chains = 4,
  iter = 4000,
  warmup = 1500,
  cores = 4,
  seed = 123
)

```

```{r}
summary(model2)
```

patch_number_z = 1.46- Participants clearly accumulate more reward over successive patches.
Environment 2 has ~83% lower cumulative reward than Environment 1.
patch_number_z:stim_catpre = -0.30- The learning slope in pre-stim is 26% flatter than in post-stim
Main anxiety effect: -0.01 -No overall anxiety effect on reward level
stim × anxiety: -0.06 -Higher anxiety people have slightly lower cumulative reward in pre condition.

patch_number_z:stim_catpre:trait_anxiety_score_z = -0.05, CI excludes zero.
In pre-stim condition, higher anxiety reduces learning slope. In post-stim condition, this anxiety-related reduction is attenuated.
So stimulation may buffer anxiety-related learning deficits.

Random slope correlation = 0.95 -Participants who start with higher reward also show steeper learning slopes
```{r}
 pp_check(model2)
```

